Preparing your manuscript

### 2025-02-12 article/article.qmd
Change small details in figures

```{r}
#| label: fig-trend
#| fig-width: 8
#| fig-height: 4
#| fig.cap: "a) Multimodel change in the frequency of $\\Delta T$ (days/year, SSP5-8.5 2018-2100 minus Historical 1980-2000) and zonal mean change over sea b) and land c), each grey line represents the models listed in Table 1 and in black, the multimodel mean. High topography areas are gray-shaded. Dotted areas show where at least 70% of the models agree in the sign of change"

@@

map_trend <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "diff") |> 
  _[, .(mean_dt = mean(deltap97.5, na.rm = TRUE)), by = .(lon, lat)] |> 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_contour_fill(aes(z = mean_dt, fill = after_stat(level)),
                    breaks = c(-Inf, seq(-8, 8, 0.25), Inf)) +
  scale_fill_divergent_discretised(name = "days/year", mid = "#f7f7f7",
                                   guide = guide_colorbar(barheight = 0.5),
                                   labels = function(x) JumpBy(x, 2, fill = "")) +
  geom_sf(data = global_map, inherit.aes = FALSE, fill = NA, linewidth = 0.2) +
  geom_point(data = sig[is.cross(lon, lat, skip = 1) & abs(sig) >= 7], alpha = 0.3, shape = 4, size = 0.1) +
  ggnewscale::new_scale_fill() +
  geom_contour_fill(data = topo, aes(ConvertLongitude(lon), lat, z = topo), breaks = c(1300, Inf), fill = "grey90") +
  coord_sf(expand = FALSE) +
  labs(x = NULL, y = NULL, fill = "days/year") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
        legend.key.width = grid::unit(1, 'null'),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), plot.margin = unit(c(0, 0 , 0 , 0), units = "mm"))

@@

model_spread <- data |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, let(serie = fcase(lat %between% c(48, 50) & land == "Sea", "50 - NH",
                        lat %between% c(38, 40) & land == "Sea", "40 - NH",
                        lat %between% c(28, 30) & land == "Sea", "30 - NH", 
                        lat %between% c(48, 50) & land == "Land", "50 - NH",
                        lat %between% c(38, 40) & land == "Land", "40 - NH",
                        lat %between% c(28, 30) & land == "Land", "30 - NH",
                        lat %between% c(-50, -48) & land == "Land", "50 - SH",
                        lat %between% c(-40, -38) & land == "Land", "40 - SH",
                        lat %between% c(-30, -28) & land == "Land", "30 - SH",
                        lat %between% c(-50, -48) & land == "Sea", "50 - SH",
                        lat %between% c(-40, -38) & land == "Sea", "40 - SH",
                        lat %between% c(-30, -28) & land == "Sea", "30 - SH"))] |>
  _[!is.na(serie) & var == "deltap97.5" & model %in% models_in_gadi] |>
  _[, .(mean_freq = weighted.mean(ssp585, cos(lat*pi/180))), by = .(serie, land, model)] |> 
  tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  _[, let(scenario = "SSP5-8.5")]

serie_regions <- serie |>
  _[!(year == 2014 & scenario != "historical")] |>
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, let(serie = fcase(lat %between% c(48, 50) & land == "Sea", "50 - NH",
                        lat %between% c(38, 40) & land == "Sea", "40 - NH",
                        lat %between% c(28, 30) & land == "Sea", "30 - NH", 
                        lat %between% c(48, 50) & land == "Land", "50 - NH",
                        lat %between% c(38, 40) & land == "Land", "40 - NH",
                        lat %between% c(28, 30) & land == "Land", "30 - NH",
                        lat %between% c(-50, -48) & land == "Land", "50 - SH",
                        lat %between% c(-40, -38) & land == "Land", "40 - SH",
                        lat %between% c(-30, -28) & land == "Land", "30 - SH",
                        lat %between% c(-50, -48) & land == "Sea", "50 - SH",
                        lat %between% c(-40, -38) & land == "Sea", "40 - SH",
                        lat %between% c(-30, -28) & land == "Sea", "30 - SH"))] |>
  _[!is.na(serie)] |> 
  _[, .(mean_freq = weighted.mean(mean_freq, cos(lat*pi/180))), by = .(land, serie, year, scenario)]

serie_regions |> 
  tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  ggplot(aes(year, mean_freq, color = lat, group = interaction(scenario, lat))) +
  geom_line(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_ysideboxplot(data = model_spread, aes(x = lat), orientation = "x") +
  scale_color_manual(values = c("#004C73", "#CC3311", "#ee8a33")) + 
  facet_grid(land~hemisphere, scale = "free") +
  labs(x = "Year", y = latex2exp::TeX("$\\Delta T$ frequency (days/year)"), color = "Latitude") +
  theme_minimal() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.8, 0.93),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "#fbfbfb", color = "#fbfbfb"),
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text = element_text(size = 12),
        ggside.axis.text.x = element_blank())

@@

```{r}
#| label: fig-eke
#| fig-width: 8
#| fig-height: 4
#| fig.cap: "Mean change in EKE and $\\Delta T$ frequency for SSP5-8.5 (2080-2100) compared with the historical period (1980-2000) for each model. Values are calculated for the northern (25$^{\\circ}$ to 55$^{\\circ}$) and southern (-25$^{\\circ}$ to -55$^{\\circ}$) hemispheres and over sea (circles) and land (triangles). Each solid line corresponds to a linear regression for each hemisphere. b) and c) show the zonal mean for EKE ($m^2 s^2$) and $\\Delta T$ (days/year) over land and sea respectively"

@@

  _[, let(land = MaskLand(lon, lat),
          region = fcase(lat %between% c(25, 55), "NH",
                         lat %between% c(-55, -25), "SH",
                         default = NA))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  na.omit() |> 
  _[, .(mean_eke = mean(eke),
        mean_deltat = mean(deltap97.5)), by = .(region, land, model)] |> 
  ggplot(aes((mean_eke - 1)*100, (mean_deltat - 1)*100)) +
  geom_point(aes(color = region, shape = land), size = 2) +
  geom_smooth(aes(colour = region), method = "lm", se = FALSE) +
  scale_color_manual(values = c("cyan4", "orange")) +
  labs(x = latex2exp::TeX("EKE change ($m^2 s^2$)"), y = latex2exp::TeX("$\\Delta T$ frequency change (days/year)"),
       color = NULL, shape = NULL) +
  theme_minimal() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.85, 0.1),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "#fbfbfb", color = "#fbfbfb"),
        legend.margin = margin(0, 0, 0, 0, unit = "mm"),
        panel.background = element_rect(fill = "#fbfbfb", color = NA))

@@

mean_trend_dt <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "diff") |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, .(mean_eke = mean(eke, na.rm = TRUE)), by = .(lat, land, model)] |> 
  ggplot(aes(mean_eke-1, lat)) +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_vline(xintercept = 0, color = "grey80") +
  geom_path(aes(group = model), color = "grey70", alpha = 0.7) +
  stat_summary(orientation = "y", geom = "line", aes(color = "EKE")) +
  geom_path(data = deltat_mean, aes(x = mean_dt, color = "DeltaT")) +
  scale_color_manual(values = c("DeltaT" = "red4", "EKE" = "black"),
                     labels = c(latex2exp::TeX("$\\Delta T$"), "EKE")) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "#fbfbfb",
            data = annotation,
            inherit.aes = FALSE) +
  scale_y_latitude(breaks = seq(-80, 80 , 20)) +
  facet_wrap(~land) +
  tagger::tag_facets(tag_pool = c("b", "c")) +
  labs(x = "Change", color = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text.x = element_text(size = 12),
        plot.margin = unit(c(0, 0 , 0, -0.40), units = "cm"),
        tagger.panel.tag.text = element_text(size = 12),
        tagger.panel.tag.background = element_rect(fill = "white", color = "white"))

>>>>

```{r}
#| label: fig-trend
#| fig-width: 8
#| fig-height: 4
#| fig.cap: "a) Multimodel change in the frequency of $\\Delta T$ (days/year, SSP5-8.5 2018-2100 minus Historical 1980-2000) and zonal mean change over sea b) and land c), each grey line represents the models listed in Table 1 and in black, the multimodel mean. High topography areas are gray-shaded. Dotted areas show where at least 90% of the models agree in the sign of change"

@@

map_trend <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "diff") |> 
  _[, .(mean_dt = mean(deltap97.5, na.rm = TRUE)), by = .(lon, lat)] |> 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_contour_fill(aes(z = mean_dt, fill = after_stat(level)),
                    breaks = c(-Inf, seq(-8, 8, 0.5), Inf)) +
  scale_fill_divergent_discretised(name = "days/year", mid = "#f7f7f7",
                                   guide = guide_colorbar(barheight = 0.5),
                                   labels = function(x) JumpBy(x, 2, fill = "")) +
  geom_sf(data = global_map, inherit.aes = FALSE, fill = NA, linewidth = 0.2) +
  geom_point(data = sig[is.cross(lon, lat, skip = 1) & abs(sig) >= 9], alpha = 0.3, shape = 4, size = 0.1) +
  ggnewscale::new_scale_fill() +
  geom_contour_fill(data = topo, aes(ConvertLongitude(lon), lat, z = topo), breaks = c(1300, Inf), fill = "grey90") +
  coord_sf(expand = FALSE) +
  labs(x = NULL, y = NULL, fill = "days/year") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
        legend.key.width = grid::unit(1, 'null'),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), 
        plot.margin = unit(c(0, -0.4 , 0 , 0), units = "cm"))

@@

model_spread <- data |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, let(serie = fcase(lat %between% c(48, 50), "50 - NH",
                        lat %between% c(38, 40), "40 - NH",
                        lat %between% c(28, 30), "30 - NH", 
                        lat %between% c(-50, -48), "50 - SH",
                        lat %between% c(-40, -38), "40 - SH",
                        lat %between% c(-30, -28), "30 - SH"))] |>
  _[!is.na(serie) & var == "deltap97.5" & model %in% models_in_gadi] |>
  _[, .(mean_freq = weighted.mean(ssp585, cos(lat*pi/180))), by = .(serie, land, model)] |> 
  tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  _[, let(scenario = "SSP5-8.5")] |> 
  _[!(land == "Land" & hemisphere == "SH" & lat == 50)]

serie_regions <- serie |>
  _[!(year == 2014 & scenario != "historical")] |>
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, let(serie = fcase(lat %between% c(48, 50), "50 - NH",
                        lat %between% c(38, 40), "40 - NH",
                        lat %between% c(28, 30), "30 - NH", 
                        lat %between% c(-50, -48), "50 - SH",
                        lat %between% c(-40, -38), "40 - SH",
                        lat %between% c(-30, -28), "30 - SH"))] |>
  _[!is.na(serie)] |> 
  _[, .(mean_freq = weighted.mean(mean_freq, cos(lat*pi/180))), by = .(land, serie, year, scenario)] |> 
  _[!(land == "Land" & serie == "50 - SH")]

serie_regions |> 
  tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  _[year >= 1900] |> 
  ggplot(aes(year, mean_freq, color = lat, group = interaction(scenario, lat))) +
  geom_line(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_ysideboxplot(data = model_spread, aes(x = lat), orientation = "x") +
  scale_color_manual(values = c("#004C73", "#CC3311", "#ee8a33")) + 
  facet_grid(land~hemisphere, scale = "free") +
  labs(x = "Year", y = latex2exp::TeX("$\\Delta T$ frequency (days/year)"), color = "Latitude") +
  theme_minimal() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.78, 0.93),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "#fbfbfb", color = "#fbfbfb"),
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text = element_text(size = 12),
        ggside.axis.text.x = element_blank(),
        ggside.panel.scale = 0.2)

@@

```{r}
#| label: fig-eke
#| fig-width: 8
#| fig-height: 4
#| fig.cap: "Mean change in EKE and $\\Delta T$ frequency for SSP5-8.5 (2080-2100) compared with the historical period (1980-2000) for each model. Values are calculated for the northern (25$^{\\circ}$ to 65$^{\\circ}$) and southern (-25$^{\\circ}$ to -65$^{\\circ}$) hemispheres and over sea (circles) and land (triangles). Each solid line corresponds to a linear regression for each hemisphere. b) and c) show the zonal mean for EKE ($m^2 s^{-2}$) and $\\Delta T$ (days/year) over land and sea respectively"

@@

  _[, let(land = MaskLand(lon, lat),
          region = fcase(lat %between% c(25, 65), "NH",
                         lat %between% c(-65, -25), "SH",
                         default = NA))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  na.omit() |> 
  _[, .(mean_eke = mean(eke),
        mean_deltat = mean(deltap97.5)), by = .(region, land, model)] |> 
  ggplot(aes(mean_eke, mean_deltat)) +
  geom_point(aes(color = region, shape = land), size = 2) +
  geom_smooth(aes(colour = region), method = "lm", se = FALSE) +
  scale_color_manual(values = c("cyan4", "orange")) +
  labs(x = latex2exp::TeX("EKE change ($m^2 s^{-2}$)"), y = latex2exp::TeX("$\\Delta T$ frequency change (days/year)"),
       color = NULL, shape = NULL) +
  theme_minimal() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.83, 0.08),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "#fbfbfb", color = "#fbfbfb"),
        legend.margin = margin(-3, 0, 0, 0, unit = "mm"),
        panel.background = element_rect(fill = "#fbfbfb", color = NA))

@@

mean_trend_dt <- data |> 
  _[(model %in% models_in_gadi) & is.finite(change)] |>
  dcast(lat + lon + model ~ var, value.var = "diff") |> 
  _[, let(land = MaskLand(lon, lat))] |> 
  _[, let(land = factor(land, labels = c("Sea", "Land")))] |> 
  _[, .(mean_eke = mean(eke, na.rm = TRUE)), by = .(lat, land, model)] |> 
  ggplot(aes(mean_eke-1, lat)) +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_vline(xintercept = 0, color = "grey80") +
  geom_path(aes(group = model), color = "grey70", alpha = 0.7) +
  stat_summary(orientation = "y", geom = "line", aes(color = "EKE")) +
  geom_path(data = deltat_mean, aes(x = mean_dt, color = "DeltaT")) +
  scale_color_manual(values = c("DeltaT" = "red4", "EKE" = "black"),
                     labels = c(latex2exp::TeX("$\\Delta T$"), "EKE")) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "#fbfbfb",
            data = annotation,
            inherit.aes = FALSE) +
  annotate(geom = "rect", ymin = c(-90, -20, 90), ymax = c(-65, 20, 65), xmin = -Inf, xmax = Inf,
           fill = "grey93", alpha = 0.7) +
  scale_y_latitude(breaks = seq(-80, 80 , 20)) +
  facet_wrap(~land) +
  tagger::tag_facets(tag_pool = c("b", "c")) +
  labs(x = "Change", color = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text.x = element_text(size = 12),
        plot.margin = unit(c(0, 0 , 0, -0.40), units = "cm"),
        tagger.panel.tag.text = element_text(size = 12))
